\Large \bf Relationship of Weather and Maize Yields in Kenya

### 2018-11-14 writing/draft1/manuscript9.tex
adding

A number of previous studies were focused on estimation of relationship of agricultural yield and measures of precipitation and temperature in Kenya or in other countries in Sub-Saharan Africa. Common measures of precipitation and temperature are monthly averages (or monthly totals in case of precipitation) and their variances or standard deviations (\citealt{AbrahaSavage2006, LobellEtAl2008, ThorntonEtAl2009}). \cite{Adejuwon2004} focused on the Sub-Saharan West Africa and he used total rainfall for: \textit{(i)} The first month of the period from sowing to harvesting (June) \textit{(ii)}: The first two months of the period from sowing to harvesting (June and July) and \textit{(iii)}: The first three months of the period from sowing to harvesting (June, July and August). Based on his results, June and July are the most important for crop yield. Other studies are based on seasonal totals or means..(..)  Some authors have utilized simulated daily extremes, averages or daily measures of variance  (\citealt{Chipanshi2003,AbrahaSavage2006,SchulzeEtA1993}). Another measure which has been used to model relationship with maize yields in the past literature is the number of wet and dry day counts per a defined period, usually a month or season (\citealt{AbrahaSavage2006}) The definition of wet and dry days vary across the studies.

@@

@@

Therefore we opted for a different approach to gauging drought severity. In particular, we decided to utilize daily precipitation data \textcolor{red}{from...CHIRPS?? with resolution... and daily temperature data from..at resolution....}. The frequency  of the maize yield data is yearly while we have daily weather data available. Hence, the weather data need to be aggregated in order to obtain a dataset conformable with the yield data. There are many possibilities how to aggregate daily weather data. Based on the literature research in Section~\ref{Introduction} and complexity of deriving the measures, we short-listed following aggregates of the weather data:

@@

\sloppy
The estimation function can be written as \citep{Friedman2010}:

\begin{equation}\label{Lasso}
\begin{array}{lcll}

 \underset{(\beta_0, \beta) \in \mathbb{R}^{(p+1)}}{min} \boldsymbol{R}_{\lambda}(\beta_0, \beta)&=&
   \underset{(\beta_0, \beta) \in \mathbb{R}^{(p+1)}}{min} \bigg[ \frac{1}{2N}\sum_{i=1}^{N}(y_i - \beta_0-x_i^{\intercal}\beta)^2 + \lambda\sum_{j=1}^{p}(|\beta_j|) \bigg], \\
\end{array}
\end{equation}

In line with common practice, we compute estimator~\eqref{Lasso} for a series of~$\lambda$ and then we choose a preferred value of~$\lambda$ using cross-validation \citep{statisHighDimData}.

\begin{equation}\label{pvals}
Q_j(\gamma)=min \left \{1,q_{\gamma} (   \{ P^{(h)}_j/\gamma; h=1,...,H\}) \right \},
\end{equation}

 where and $q_{\gamma}(\tiny{\cdot})$ is the (empirical) $\gamma$-quantile function. We will further refer to this procedure as a multisplit lasso.

@@

{\centering
\begin{threeparttable}
\caption{\textit{\textbf{Sex and age distribution} \\ of the sample and the population}}
\label{SexAge} 
\begin{small}
\begin{tabular}{|l|rr|rr|} 
\hline	
  \multicolumn{1}{|l|}{ } & \multicolumn{2}{c}{\bf{Sample}} & \multicolumn{2}{|c|}{\bf{UK population\tnote{a}}} \\ %sample pubpolM2 used here
    \multicolumn{1}{|l|}{\vspace{0.1cm}\textbf{Age range}}  &  \multicolumn{1}{|c}{\bf{Male}} & \multicolumn{1}{c}{\bf{Female}} & \multicolumn{1}{|c}{\bf{Male}} & \multicolumn{1}{c|}{\bf{Female}}\\ 
\hline 
$18-24$&$9.8\%$&$9.4\%$&$6.2\%$&$6.0\%$\\
$25-34$&$10.0\%$&$10.3\%$&$9.0\%$&$9.1\%$\\
$35-44$&$7.8\%$&$8.3\%$&$8.6\%$&$8.8\%$\\
$45-54$&$8.1\%$&$9.4\%$&$9.3\%$&$9.6\%$\\
$55-64$&$7.4\%$&$8.4\%$&$7.5\%$&$7.8\%$\\
$65-74$&$4.1\%$&$4.8\%$&$6.2\%$&$6.7\%$\\
$75-80$&$0.1\%$&$0.1\%$&$2.4\%$&$2.8\%$\\
\hline
\hline
\end{tabular} 
  \begin{tablenotes}
  \begin{footnotesize}
  \singlespacing
     \item[a]Population data are from the Office of National Statistics, Population Estimates of UK, England and
Wales, Scotland and Northern Ireland Mid $2014$, Table MYE2.
\singlespacing
  \end{footnotesize}
\end{tablenotes}
\end{small}
  \end{threeparttable} \par}

\hspace{1.5cm}

@@

{\centering
\begin{threeparttable}
\caption{\textit{\textbf{Dependent variables:} Descriptive statistics}}
\label{DepDes} 
\centering
\begin{small}
\begin{tabular}{lrrrr} 
\hline	
  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable:}}  &  \multicolumn{1}{c}{\bf{Mean}} & \multicolumn{1}{c}{\bf{St. dev.}} & \textbf{Min} & \textbf{Max}\\ 
\hline \vspace{-0.3cm} \\ 
  \vspace{0.15cm}Climate change knowledge&$3.851$&$1.266$&$1$&$8$\\
    \vspace{0.15cm}Climate change seriousness perception&$6.622$&$2.249$&$0$&$10$\\
        \vspace{0.15cm}Climate versus policy effects perception&$5.370$&$2.315$&$0$&$10$\\
\vspace{0.15cm}WTP - gas and electricity tax (\textsterling~per year)&$123.900$&$105.459$&$0$&$500$\\
 \vspace{0.15cm}WTP - duty on transport fuel (pence per year)&$20.530$&$22.518$&$0$&$100$\\
\hline
\hline
\end{tabular} 
\end{small}
  \end{threeparttable} 
\par}

\hspace{1.5cm}

@@

To estimate the social value orientation, respondents played six dictator games with the same questions as in \citet{murphy2011SVO}. The ring measure of social value orientation which we use in our models is defined as

\begin{equation}\label{RingMeasure}
\begin{array}{lcll}

R&=&\arctan \frac{\sum_{i=1}^{N} P_O -50N}{\sum_{i=1}^{N} P_S -50N}, \\
\end{array}
\end{equation}

\FloatBarrier
	\section{Results and discussion}\label{Results}
In this section we describe our results and discuss their interpretation.

In the tables which summarise the estimates of lasso below, \textit{p}-values of some of the explanatory variables are equal to one. These variables were not selected by the lasso in most of the sample splits. They are, however, included in the tables because they represent either a category of a nominal variable whose other category was selected by the lasso or a linear term of a variable whose quadratic term was selected by the lasso.

>>>>

A number of previous studies were focused on estimation of relationship of agricultural yield and measures of precipitation and temperature in Kenya or in other countries in Sub-Saharan Africa. Commonly used measures of precipitation and temperature include monthly averages (or monthly totals in case of precipitation) and their variances or standard deviations (\citealt{AbrahaSavage2006, LobellEtAl2008, ThorntonEtAl2009}). \cite{Adejuwon2004} analysed relationship of crop yields and three measures of precipitation. The three measures include total rainfall during: \textit{(i)} The first month of the period from sowing to harvesting (June) \textit{(ii)} The first two months of the period from sowing to harvesting (June and July) and \textit{(iii)} The first three months of the period from sowing to harvesting (June, July and August). Based on his results, weather during June and July is the most important for crop yield in Sub-Saharan West Africa. Other studies utilised seasonal totals or means (\citealt{sagoe2006,LobellBurke2010} \textcolor{red}{and others,find!!})  Some authors have utilized simulated daily extremes, averages or daily measures of variance  (\citealt{SchulzeEtA1993,Chipanshi2003,AbrahaSavage2006}). Another measures which have been proposed for modelling the variability of maize yield are numbers of wet and dry days per a defined period, usually a month or a season (\citealt{BenMohamed2002,AbrahaSavage2006,sagoe2006,Giannakopoulos2009})   or length of rainy season (\citealt{Leemans1993,BenMohamed2002}). The definition of wet and dry days and rainy season vary across the literature. For example, \citep{BenMohamed2002} has assumed that rainy season begins when the amount of rainfall in three consecutive days is at least 25mm and no dry spell of more than seven days duration occurs in the following thirty days and the end of the rainy season is that rainy day after which rain recorded during 20 days is less than 5mm.

@@

\subsection{Measures of Yield and climate}

@@

Therefore we opted for a different approach to gauging drought severity. In particular, we decided to utilize daily precipitation data \textcolor{red}{from...CHIRPS?? with resolution... and daily temperature data from..at resolution....}. The frequency  of the maize yield data is yearly while we have daily weather data available. Hence, the weather data need to be aggregated in order to obtain a dataset conformable with the yield data. There are many possibilities how to aggregate daily weather data. Based on the literature research in Section~\ref{Introduction} and complexity of deriving the measures, we short-listed the following aggregates of the weather data:

@@

\subsection{Linear Mixed Models}

\sloppy

@@

@@

@@

\FloatBarrier
	\section{Results and discussion}\label{Results}
